Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-16 paper/high-assurance-contracts.tex
Complete the AGREE explanations in section 3

    -- start user definitions
    const MAX_LATENCY : int = 1;

@@

    eq isPending : bool = Since(not rsp, req and not rsp);
    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);

    eq policy : bool = (rsp => req) ->
                        (    (isPending => latency < MAX_LATENCY)   
                        and (rsp => (req or pre(isPending))));
    -- stop user definitions

    eq alert : bool = (not policy) -> 
                    ((is_latched and pre(alert)) or not policy);

    assume "One outstanding request at a time" :
    (true -> (req => not pre(isPending)));

@@

As noted previously, the system with the untrusted AI does not pass AGREE verification.
Transformations in BriefCASE are used to cyber-harden the system against malicious AI behavior by inserting a high-assurance filter and monitor between the AI and the WM as shown in \figref{fig:hardened}.
High-assurance means the behaviors of the components are verified.
Here the filter protects the WM from malformed data and the monitor to protects it from spontaneous or delayed responses.
These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear, and then choosing the appropriate transformation.

The system designer provides configuration parameters for the transformation in a dialogue box as shown in \figref{fig:dialogue} for the filter.
All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
A filter policy defines well-formed data while a monitor policy defines an invariant over inputs and outputs.
These policies can be stated directly in the wizard, or they can be left blank and added later to the AGREE contract generated by the transformation.

The BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
It also creates default AGREE specification that define everything about the component behavior but the policy itself unless it is stated in the dialogue.
The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
Here the policy \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)} was specified in the dialogue box.
The single guarantee completely defines the filter output under any input according to the truth value of the policy.
While such a string guarantee is not required for AGREE analysis, it is required for automated synthesis.

The resulting AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerable more complex than the filter.

The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
This behavior is one of the several options available in the dialogue.
The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

>>>>

    -- start user definitions
    const MAX_LATENCY : int = 1; *\label{line:mon-const}*

@@

    eq isPending : bool = Since(not rsp, req and not rsp);*\label{line:mon-pending}*
    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);*\label{line:mon-latency}*

    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
                         (    (isPending => latency < MAX_LATENCY)   
                          and (rsp => (req or pre(isPending))));

    assume "One outstanding request at a time" : *\label{line:mon-assume}*
      (true -> (req => not pre(isPending))); 
    -- stop user definitions

    eq alert : bool = (not policy) -> *\label{line:mon-alert}*
                        ((is_latched and pre(alert)) or not policy);

@@

As noted previously, the system with the untrusted AI does not pass AGREE verification.
Transformations in BriefCASE are used to cyber-harden the system against malicious AI behavior by inserting a high-assurance filter and monitor between the AI and the WM as shown in \figref{fig:hardened}.
High-assurance means the behaviors of the components are verified.
Here the filter protects the WM from malformed data and the monitor protects it from spontaneous or delayed responses.
These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear and then choosing the appropriate transformation.

The system designer provides configuration parameters for the transformation in a dialogue box.
The dialogue box for adding a filter is shown in \figref{fig:dialogue}.
All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
A filter policy defines well-formed data while a monitor policy defines an invariant over inputs and outputs.
These policies can be stated directly in the wizard, or they can be left blank and added later to the AGREE contract generated by the transformation.

BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
It also creates a default AGREE specification that defines everything about the component behavior but the policy itself unless it is provided in the dialogue.
The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
The single guarantee completely defines the filter output under any input according to the truth value of the policy.
While such a strong guarantee is not required for AGREE analysis, it is required for automated synthesis.

The AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerably more complex than the filter.
BriefCASE generates everything but the statements in \linesref{line:mon-const}{line:mon-assume}.
As with the filter, the guarantees completely define the output, as required for synthesis, in terms of the policy and the alert.
The \texttt{alert} variable defined on \lineref{line:mon-alert} depends not just on the policy but also on the indicated behavior from the BriefCASE dialogue box. 
Marking \texttt{is\_latched} as true makes the alert persistent, meaning that once the alert is raised, it is always raised; otherwise, it is the complement of the policy value in the current step.

The monitor policy is defined by marking when a request is not satisfied, \texttt{isPending} on \lineref{line:mon-pending}, and counting the number of steps between requests, \texttt{latency} on \lineref{line:mon-latency}.
The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response--\emph{not response since request and not response}.
The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.

The latency bound for the monitor is defined by the constant \texttt{MAX\_LATENCY} on \lineref{line:mon-const}.
Here the constant is set to one to be consistent with the system specification from the previous section.
At time 0 the policy is trivial: a response requires an accompanying request.
After time 0, if a request is pending in the current time step, then the latency must still be under the bound, and if there is a response, then it closes a request now or a pending request from earlier in time.

The policy definition only works if there is never more than one outstanding request at a time.
That requirement is encoded in the assumption on \lineref{line:mon-assume}.
A request is only allowed if one is not already pending in the pre-state of the component.
The policy can be written without needing the assumption, but it is much simpler to write with the assumption.

As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
Synthesis must generate precise implementations of the high-assurance component specifications for the AGREE verification result to apply to the deployed system.
Precise in this context means that the implementations match the input to output relations defined in the specifications.

### 2022-03-16 paper/paper.tex
Complete the AGREE explanations in section 3

\lstdefinestyle{agree}
{frame=none,
  basicstyle=\ttfamily,
  language=ML,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  numbers=none,
  numberstyle=\tiny\color{gray},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=false,
  breakatwhitespace=true,
  tabsize=2,
  linewidth=2\linewidth,
  morecomment=[l]{--},
  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
}

@@

\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Sec.~\ref{#1}}
\newcommand{\lineref}[1]{Line~\ref{#1}}
\newcommand{\linesref}[2]{Lines~\ref{#1}-\ref{#2}}

@@

\subsection{Revisions}
\begin{compactitem}
  \item Add in the AGREE specification for the system
  \item Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)
\end{compactitem}

>>>>

\lstdefinestyle{agree}
{frame=none,
  basicstyle=\ttfamily,
  language=ML,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  numbers=none,
  numberstyle=\tiny\color{gray},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=false,
  breakatwhitespace=true,
  tabsize=2,
  linewidth=2\linewidth,
  morecomment=[l]{--},
  escapeinside={*}{*},
  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
}

@@

\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Sec.~\ref{#1}}
\newcommand{\lineref}[1]{Line~\ref{#1}}
\newcommand{\linesref}[2]{Lines~\ref{#1}--\ref{#2}}

@@

\subsection{Revisions}
\begin{compactitem}
  \item \sout{Add in the AGREE specification for the system}
  \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
\end{compactitem}
